---
title: "Tidy Tuesday"
author: "Rebekah Jacob"
date: "7/21/2020"
output: html_document
---

# Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdark)
library(extrafont)
library(ggsci)
library(DT)
```

# Data load
```{r message=FALSE, warning=FALSE}
brisbane_complaints <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-21/brisbane_complaints.csv')
```

# Data management
Want to look at complaints over time, but the dates are messy and need cleaning.
```{r}
datatable(brisbane_complaints %>%
  group_by(date_range) %>%
  tally())
```

I create a "months" vector to pass to str_extract, then extract year and month name from string.
```{r}
months<- paste(c(tolower(month.abb), tolower(month.name)), collapse = "|")

df<- brisbane_complaints %>%
  mutate(year=str_extract(date_range, "[0-9]{4}"),
         months=str_extract(date_range, months))
```

You can see that one won't work because it only says 1st quarter, not a month date.
```{r}
datatable(df %>%
  group_by(year, months, date_range) %>%
  tally())
```

So now I mutate months to a quarter number (1 through 4) from month name (so I can paste to year in next step). I know the one category is actually from Q1, so I use TRUE~"1" to force that in since there were no other NA's.
```{r}
df<- df %>%
  mutate(
    qtr=case_when(
     months=="jan"~"1",
     months=="apr"~"2",
     months=="jul"~"3",
     months=="oct"~"4",
     TRUE~"1"),
    yr_qtr=paste(year, qtr, sep = "-"))
```

Now I just want to grab the most common categories of complaints to make the plot readable. So I look to see which one's have the most complaints total. Then I create a new df with just those catagories.
```{r}
datatable(df %>%
  group_by(category) %>%
  tally() %>%
  arrange(desc(n)))

most_common<- df %>%
  filter(category %in% c("Fencing Issues", "Wandering", "Attack On An Animal", "Attack On A Person"))
```

# Theme
```{r}
my_theme2<- dark_theme_gray(base_family = "Rockwell", base_size = 14) + 
  theme(plot.title = element_text(family = "Rockwell", hjust=0.5),
        plot.background = element_rect(fill = "grey10"),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey30", size = 0.2),
        panel.grid.minor = element_line(color = "grey30", size = 0.2),
        legend.background = element_blank(),
         legend.position = "bottom", 
        axis.text.x = element_text(angle=45),
        legend.key = element_blank(),
        legend.title = element_blank())
```

# Plot
```{r}
most_common %>%
  group_by(yr_qtr, category) %>%
  tally() %>%
  ggplot(aes(x=yr_qtr, y=n, group=category, color=category)) + geom_line(size = 1.3) +
  my_theme2 +
  scale_color_locuszoom() +
  labs(title = "Brisbane's most common animal complaints \nby year and quarter",
             x = "", y = "Number of complaints", legend = "",
             caption = "@RRJacob_STL")
```

# Save
```{r}
ggsave(file="Aussie_animal_complaints.png", width=10, height=7)
```

